Understanding the Nanoscale Structure of Inverted Hexagonal Phase Lyotropic Liquid Crystal Polymer Membranes

### 2018-11-19 Ben_Manuscripts/structure_paper/response_to_reviewers.tex

\begin{enumerate}

    \item \begin{quote} \textit{This manuscript describes atomistic molecular
	    dynamics simulations and diffraction analysis of hexagonally-arranged pores,
	    each of which is made of many columns of the monomer Na-GA3C11. This physical
	    system is of interest because such nanopores have potential to transform
	    filtration applications. This work is of potential interest to the JPC
	    readership because it provides atomistic insight into the origins of
	    diffraction reflections observed in experiments, which could be used to resolve
	    discrepancies about what the structure of these hexagonal arrays are, or used
	    to engineer better filters. Overall, the manuscript provides comprehensive
	    descriptions of metrics used to evaluate the simulated structures against
	    experiments, but the number of things described and length of these
	    descriptions obscures whether significant new physical insight into this system
	    is made.} 
    \end{quote}

    Author reply: We thank the reviewer for their constructive feedback and aim to 
    address all of their concerns.

    \item \begin{quote}
    \textit{As written, it is not clear if the authors are presenting this
	    study as a validation of GAFF and the present equilibration techniques, or
	    whether this study is providing significant new physical insight into these
	    lyotropic liquid crystals. The paper reads a bit more like the former, with a
	    comprehensive discussion of the degree to which five scattering features
	    expected in experiments are predicted in simulations. However, at the same
	    time, the authors make arguments including (1) pores are made from five columns
	    of monomers, (2) that water is present in the ``dry" structures, and (3) that a
	    scattering feature comes from tail packing, not tilt: These observations
	    *could* be new physical insight into lyotropic liquid crystal structure, but as
	    presented seem more like hypotheses than conclusions:}

@@

    \end{quote}

    Author reply: We thank the reviewer for offering their interpretation of
    the purpose of our study, as it is important that we properly convey our
    message. The main purpose of our work is to give a detailed description of the
    structure of a lyotropic liquid crystal membrane backed by experimental data.
    This work lays a foundation for future transport studies. By matching 2D-WAXS
    data, we have attempted to unveil the most experimentally-consistent structure
    possible. To clarify this purpose, we have modified and added text to the
    paragraph on page~\pageref{M-rework:purpose} of the introduction, in which we
    state our perceived importance of knowing the most experimentally-consistent
    structure:

@@

    Old text: ``A molecular-level understanding of LLC polymer membrane
    structure, enabled by molecular dynamics (MD) simulations, can provide
    guidelines to reduce the large chemical space available to design monomers for
    creation of separation-specific membranes. Useful molecular-level modeling
    should incorporate a detailed picture of the nanoscopic pore structure, which
    is crucial to understanding the role of monomer structure in solute transport
    and membrane design.

    New text: ``A molecular-level understanding of LLC polymer membrane
    structure, enabled by molecular dynamics (MD) simulations, will enhance our
    comprehension of the role of monomer structure in solute transport and provide
    guidelines to reduce the large chemical space available to design monomers for
    creation of separation-specific membranes. It is important that we model the
    system with a nanoscopic pore structure that is maximally consistent with
    experimental data in order to best approximate the chemical environment experienced
    by solutes in future simulated transport studies."

%This is unlikely to be achieved if one assembled these systems using
%    united atom or coarse grained models due to the large number of potential
%    metastable configurational basins which plague a system like ours which exhibts
%    such slow dynamics.

    We do not intend for this study to be a validation of GAFF for this
    particular system. We believe we would come to similar conclusions using other
    force fields, however validation of this hypothesis is left for a future study.
    We chose GAFF because it is widely used to simulate organic molecules and there
    are no catch-all force fields built for liquid crystal systems. We are interested in
    what information we can gain from a widely-used `out-of-the-box' force field and in future
    work, we will validate whether GAFF is sufficient for reproducing trends in
    experimental separation properties. To clarify our use of GAFF, we have added
    the following text to Secton~\ref{M-method:parameterization} of the Methods on
    page~\pageref{M-method:parameterization}.

    New text: ``We chose GAFF because it has been parameterized for use with
    organic molecules. Although researchers have optimized force fields, including
    GAFF, for use with liquid crystals, they are often tuned in order to
    reproduce one or more experimental observables for specific types of liquid 
    crystal molecules.$^{29,30}$ Since liquid crystals exist in a wide range of shapes and
    sizes, including candidate monomers which we may study, we chose to stick with
    a widely used force field. If we can reproduce experimental trends using a 
    single `out-of-the-box' force field, then there will be no need to complicate
    our studies with customized parameter sets."

    We have added the following citations:
    \begin{itemize}
    \item Cook, M. J.; Wilson, M. R. Development of an All-Atom Force Field for the Simulation
    of Liquid Crystal Molecules in Condensed Phases (LCFF). Mol. Cryst. Liq. Cryst.
    2001, 357, 149–165.
    \item Boyd, N. J.; Wilson, M. R. Optimization of the Gaff Force Field to Describe Liquid
    Crystal Molecules: The Path to a Dramatic Improvement in Transition Temperature
    Predictions. Phys. Chem. Chem. Phys. 2015, 17, 24851–24865.
    \end{itemize}

    Since our ultimate goal is to design new monomers for solute-specific
    separations, we will be employing the same techniques used in this work to
    unsynthesized monomers. In most cases, we will not have an experimental dataset
    to use for comparison. Despite all of the structural intricacies which we have
    revealed in this study, one of our main conclusions is that regardless of which
    system we study, the radial distribution of monomer components stays
    consistent. From this, we can hypothesize that we will extract similar
    transport results regardless of which configuration we use to conduct
    transport studies. This is an extremely important finding which may have
    been undersold in the original draft of the manuscript. We have added the
    following concluding paragraph to the end of
    Section~\ref{M-section:pore_composition} on
    page~\pageref{M-addition:pore_composition} in order to drive the point home:

@@

    New text: ``Our observations suggest that the details of transport may be
    relatively independent of the structural differences between different possible
    structures studied here. Despite the structural intricacies that give rise to
    differences between the various metastable basins, all pores are characterized
    by dense, primarily hydrophilic cores with a gradient towards a primarily
    hydrophobic tail region far from the pore center. This implies that we might
    see the same trends in transport properties from any of the systems studied so
    far. Since the ultimate goal of our work is to design new monomers using high
    throughput simulations, achieving that goal becomes tractable without the need
    to optimize every system with an experimental dataset."

    We also modified a paragraph in our conclusion on page~\pageref{M-rework:conclusion}
    to better emphasize our point:

@@

    New text: ``Although exactly reproducing the experimental 2D-WAXS pattern
    required attention to minute structural details, we found that all systems
    showed similar radial distribution functions characterized by a gradual radial
    transition from a dense hydrophlic core to hydrophobic tails. This observation
    enables us to study new systems that have not been experimentally characterized
    with the expectation that any additional structural optimization will not
    greatly influence trends in transport property predictions. The compositional
    gradient itself raises questions about the nature of size-exclusion separations
    in systems without well-defined pore boundaries, which potentially could enable
    separations that vary with chemical identity as well as size."

    \item \begin{quote}
    \textit{1. For the simulations performed here, the authors do make a strong
	    case that 5 columns per core gets the best pore-to-pore spacing (Figure 5).
	    However, insufficient time is accessed to see whether the 4-, 5-, and 6-column
	    initial conditions move towards or away from the expected pore spacing over
	    many-microseconds trajectories. That said, for the quenches performed here,
	    5-columns per core seems more plausible than 4.}
    \end{quote}

    Author reply: We agree with the reviewer that 5 columns-per-core is more
    plausible than 4 and that much longer simulations may provide further
    supporting evidence. However, there is very little motion on the hundreds of
    nanoseconds timescale that we simulated which suggests that we will not gain
    any significant new insight by extending our simulations into the microsecond
    regime. It is not feasible for us to carry out these simulations as it would
    consume our computational resources with little reward.
    %BJC: The following might apply:
    Some confusion may have been generated by agreement of the disordered basin
    6 column-per-pore system with experimental pore spacing. As mentioned in the
    text on page~\pageref{M-6colperpore}, the vertical stacking distance between 
    monomers in that system is far too large. The columns are `stretched' in the
    $z$-direction allowing the pores take up less space on the $xy$ plane and 
    thus pack closer together.

@@

    \item \begin{quote}    
    \textit{Is this significant physical insight? It is insight, but the
	    authors do not make a compelling case for how 5 columns would change how we
	    think about these crystals, or how it would impact membrane engineering (or
	    more importantly, whether these 5-column structures assemble in the solution
	    phase, or are only locally stable).}
    \end{quote}

    Author reply: We thank the reviewer for pushing us to provide more
    significant physical insight based on the number of monomers per column. The
    number of columns-per-pore influences the pore radius which is important for
    determining the limits of size-exclusion-based separations. It also tells us
    the density of interacting sites within the pore which may play a role in small
    molecule transport. We have added the following text to the main document at
    the end of Section~\ref{M-section:mon_per_pore} on page~\pageref{M-addition:pore_radius}:

    ``The number of columns per pore dictates the density of interaction sites
    within each pore and the pore radius. A higher number of interaction sites will
    surely play a role in transport of molecules that have an affinity for the
    monomer head groups.  The pore radius increases with the number of columns per
    pore (See Figure~\ref{S-fig:pore_radius}). This radius is likely related to the
    size of the molecules which can be excluded by this type of membrane. One may be
    able to control LLC membrane pore size by selecting monomers that are stable in
    configurations with more columns-per-pore."

    We are unable to definitively answer the question of whether the 5-column 
    structures are most likely to assemble in solution phase. The system does
    not self-assemble on a timescale reasonable for simulation. See 
    Section~\ref{S-section:self_assembly} of the supporting information where
    we attempted self-assembly. These sorts of predictions are left for 
    a separate effort that will use coarse grain models.

@@

    \item \begin{quote}
    \textit{2. The authors make a pretty compelling case that water causes the
    R-double scattering feature by showing its absence when water is added, but
    that the R-double appears and persists for 200ns after the water makes its way
    into the crystal. However, the authors also show that nonuniformly spaced
    monomers can give rise to the R-Double feature, a seeming counterexample to the
    necessity of water.  Is the R-double feature with water because the monomers
    are now irregularly spaced? What happens when the waters are removed and the
    scattering re-performed in Fig 17c(bottom)?}
    \end{quote}

    Author reply: We thank the reviewer for exposing a potentially unclear
    explanation of the origin of R-double. R-double appears in a dry system with
    non-uniformly spaced monomers \textit{only} if $z$-direction position
    restraints are applied to monomer head groups in order to maintain unevenly
    spaced stacking. As soon as the restraints are released, R-double fades and the
    monomer spacing becomes uniform. We modified the text on page~\pageref{M-modification:rdouble}
    in order to more clearly convey this message:

    Old text: ``We can also produce R-double if the LLC monomers are not
    uniformly spaced in the $z$-direction, but instead form pairs that stack less
    than 3.7~\AA~apart, with COMs that are spaced 7.4 \AA~ from neighboring pairs
    of monomers (Figure~\ref{M-fig:staggered_rzplot_norestraints}). Our force field
    causes our system to tend towards uniformly spaced layers. Simulations of
    unevenly spaced systems are only stable if position restraints are applied to
    heavy atoms of the phenyl rings. Additionally, there is little evidence from QM
    studies of stacked $\pi-\pi$ systems that such uneven stacking could be
    energetically stable.$^{52}$%.~\cite{tauer_beyond_2005}"

    Modified text: ``We can also produce R-double if the LLC monomers are not
    uniformly spaced in the $z$-direction, but instead are placed in pairs that
    stack less than 3.7~\AA~apart, with COMs that are spaced 7.4 \AA~ from
    neighboring pairs of monomers (Figure~\ref{M-fig:staggered_rzplot_norestraints}).
    Simulations of unevenly spaced systems are only stable while position
    restraints are applied to heavy atoms of the phenyl rings. There is little
    evidence from quantum mechanical studies of stacked $\pi-\pi$ systems that such
    uneven stacking could be energetically stable.$^{52}$ As soon %~\cite{tauer_beyond_2005} As soon
    as we remove position restraints, our system to tends towards uniformly spaced
    monomers."

    R-double is only long-term stable without position restraints if a small
    amount of water is present. We believe that water facilitates the appearance
    of R-double by stabilizing non-uniform spacing through hydrogen bonding, as the
    reviewer suggests. We clarified the text on page~\pageref{M-modification:rdouble2}
    in order to make this conclusion clear:

    Old text: ``When two vertically stacked monomer head groups hydrogen bond
    with a shared water molecule, the monomers are drawn closer together (as
    illustrated in Figure~\ref{M-fig:hbond_visualization}), which creates an
    asymmetry that allows R-double to appear. If a monomer head group shares a
    hydrogen-bonded water molecule with a head group above itself, it will be less
    likely to share a water molecule with a head group below it due to geometric
    constraints. The monomer head group below can just as easily share a water
    molecule with a head group below itself. In this scenario, the COMs of each
    pair are 2 times the $\pi$-stacking distance apart which would lead to R-double
    (much like the configuration in
     Figure~\ref{M-fig:staggered_rzplot_norestraints}). There are a modest number of
    occurrences of this scenario, which we quantify in further detail in the SI,
    Section \ref{S-section:rdouble}."

    New text: ``The addition of a small amount of water to the pores stabilizes
    non-uniform stacking of head groups much like that shown in
    Figure~\ref{M-fig:staggered_rzplot_norestraints}. When two vertically stacked
    monomer head groups hydrogen bond with a shared water molecule, the monomers
    are drawn closer together (as illustrated in
    Figure~\ref{M-fig:hbond_visualization}), which creates the asymmetry that allows
    R-double to appear. If a monomer head group shares a hydrogen-bonded water
    molecule with a head group above itself, it will be less likely to share a
    water molecule with a head group below it due to geometric constraints. The
    monomer head group below can just as easily share a water molecule with a head
    group below itself. There are a modest number of occurrences of this scenario,
    which we quantify in further detail in the SI, Section
    \ref{S-section:rdouble}."

    The reviewer asked us to re-perform the scattering analysis on the same
    system with water molecules removed. We would like to emphasize that we
    attempted to make a similar argument by removing just the head group atoms from
    the trajectory and re-performing the scattering analysis
    (Figure~\ref{M-fig:rdouble_nophenyls}). In the absence of head groups, R-double
    is not present, implying that their structure (which is influenced by the
    presence of water molecules) is responsible for the R-double reflection. For
    completeness, we performed the scattering analysis on the same system with
    water molecules removed, as suggested by the reviewer, and R-double is still
    present. This is the expected result since the structure of water itself does
    not give rise to R-double. We've added the resulting simulated XRD pattern to
    the Supporting Information, Figure~\ref{S-fig:rdouble_water_removed}.

@@

    \item \begin{quote}
    \textit{3. The authors make the case that the R-spots feature comes from the 
    structure of the hexagonal tails, not their tilt. The authors should perform the
    scattering analysis to check if the R-spots are present on the tilted-tails they
    constrain: This quick experiment would nail down their interpretation here, or 
    show that the tilted tails are also plausible (but not observed in this model).}
    \end{quote}

    Author reply: We thank the reviewer for their suggestion. We do not deny
    that the tilt of the monomer tails is a plausible explanation for the
    appearance of reflections that resemble R-spots in many liquid crystal systems.
    While the reviewer's suggested experiment is a good idea, its results in this
    case are ambiguous for a couple reasons. First, the tails may still pack
    hexagonally in the tilted configuration which would independently give rise to
    R-spots. Second, it is necessary to impose high force constant ($>$ 1000 kJ mol$^-1$
    nm$^{-2}$) position restraints in order to maintain a tilt angle close to that
    implied by the 2D WAXS data. The resulting high degree of order in the system 
    gives way to reflections which are easily misinterpretted. We have added the
    following text highlighted in red to the text shown on page~\pageref{M-addition:rspots} in order to
    acknowledge that tail tilt is a valid explanation of R-spot-like reflections:

    Modified text: ``Previous literature has attributed the R-spots reflection
    in this particular WAXS dataset as the result of tilted alkane
    chains $^{16}$. {\color{red} This explanation is not  % feng_scalable_2014
    unfounded as there are examples in literature of systems where tilted liquid
    crystals give rise to reflections that resemble R-spots.$^{45,46}$} We
    measured the tilt angle of the alkane chains ..."

    We have added the following citations in support of the above statement:
    \begin{itemize}

    \item Percec, V.; Imam, M. R.; Peterca, M.; Wilson, D. A.; Graf, R.; Spiess, H. W.; Balagu-
    rusamy, V. S. K.; Heiney, P. A. Self-Assembly of Dendronized Triphenylenes into Helical
    Pyramidal Columns and Chiral Spheres. J. Am. Chem. Soc. 2009, 131, 7662–7677.

    \item Gearba, R. I.; Anokhin, D. V.; Bondar, A. I.; Bras, W.; Jahr, M.; Lehmann, M.;
    Ivanov, D. A. Homeotropic Alignment of Columnar Liquid Crystals in Open Films by
    Means of Surface Nanopatterning. Adv. Mater. 2007, 19, 815–820.

    \end{itemize}

@@

    \item \begin{quote}
    \textit{These arguments would be stronger if there weren't so many
	    qualifications about simulation timescales and GAFF, in particular its ability
	    to predict the shortest-scale feature (pi-stacking) for this system. How does
	    everything else change if the force-field were tuned to get that feature
	    correct?}
    \end{quote}

    Author reply: We thank the reviewer for their constructive criticism. Many
    of our findings may be relatively independent of the force fields used, but we
    feel that we should include these qualifications in order to not overstate our
    case. It is common for researchers to run simulations for an insufficient
    length of time and to make bold claims. While we feel that we have run our 
    simulations sufficiently long, we still acknowledge the shortcomings of our
    techniques in order to inspire other researchers to build on our initial work.

    We agree with the reviewer that we could tune our force field in order to
    exactly reproduce the $\pi-\pi$ stacking distance seen experimentally. Surely
    there will be some structural differences. There are many ways to tune a force
    field in order to achieve this. As stated earlier, we prefer to work with an
    `out-of-the-box' force field since they are widely used and can be applied to
    vast number of systems. The ultimate goal of our work is to create new monomers
    that can achieve specfic separations. If we can gain experimentally-consistent
    insights using our model simulated with GAFF, then we can have increased
    confidence in future predictions generated from systems built with
    unsynthesized monomers.  With that said, we think that it would be interesting
    to repeat our analysis using different and/or optimized force fields in a
    future study in order to comment on the signficance of any differences.

### 2018-11-19 Ben_Manuscripts/structure_paper/Final_Draft.tex
Complete draft of response to reviewers

  % reworking paragraph
  A molecular-level understanding of LLC polymer membrane structure, enabled by
  molecular dynamics (MD) simulations, will enhance our comprehension of the role
  of monomer structure in solute transport and provide guidelines to reduce the
  large chemical space available to design monomers for creation of
  separation-specific membranes. It is important that we model the system with a
  nanoscopic pore structure that is maximally consistent with experimental data
  in order to approximate the chemical environment experienced by solutes in 
  future transport studies. Atomistic MD simulations can provide the required
  level of detail (Figure \ref{fig:detailed_pore}), assuming the force fields are
  sufficiently accurate. With such an atomistic model, we can directly observe
  molecular-level solute transport and suggest governing mechanisms. We can also
  observe how the choice of head group interacts with solutes of interest. In
  addition, we can interchange counterions which may influence both the pore size
  and the strength of the Donnan potential.

@@

  We parameterized the interaction potential for the liquid crystal monomer
  Na-GA3C11 using the Generalized AMBER Force Field (GAFF)
  \cite{wang_development_2004} with the Antechamber package
  \cite{wang_automatic_2006} provided with AmberTools16
  \cite{case_ambertools16_2016}. We chose GAFF because it has been parameterized
  for use with organic molecules. Although researchers have optimized force
  fields, including GAFF (cite), for use with liquid crystals, they are often
  tuned in order to reproduce one or more experimental observables. Since liquid
  crystals exist in a wide range of shapes and sizes, including candidate
  monomers which we may study, we chose to stick with a widely used force field.
  If we can reproduce experimental trends using a single 'out-of-the-box' force
  field, then there will be no need to complicate our studies with customized
  parameter sets. We assigned atomic charges using the am1bccsym method of
  \texttt{molcharge} shipped with QUACPAC from Openeye Scientific Software. We
  ran all molecular dynamics simulations using GROMACS 2016
  \cite{bekker_gromacs:_1993,berendsen_gromacs:_1995,van_der_spoel_gromacs:_2005,hess_gromacs_2008}.

@@

  \subsubsection{\textit{Determining equilibration time}}\label{method:equil_time}

@@

  \subsubsection{\textit{Calculation of pore spacing}}\label{method:pore_spacing}

@@

  \subsubsection{\textit{Generation of simulated X-ray diffraction patterns}}\label{method:xrd}

@@

  \subsubsection{\textit{Pair distribution functions and correlation length}}\label{section:correlation_length}

@@

  \subsubsection{\textit{Radial distribution functions}}\label{method:rdfs}

@@

  \subsection{\textit{Ionic conductivity calculations}}\label{method:ionic_conductivity}

@@

  \subsection{Density of monomers around pores}\label{section:mon_per_pore}

@@

  Systems built with 5 columns around each pore equilibrate to a pore spacing
  that is most consistent with the experimental value of 4.12 nm derived from
  SAXS measurements (Figure~\ref{fig:SAXS}). Ordered basin systems built with 4
  columns per pore equilibrate to an average pore spacing 0.25 nm lower than
  experiment. Ordered basin systems built with 6 columns per pore, have an
  equilibrated pore spacing ca. 0.50 nm higher than experiment. Monomers in
  disordered basin systems built with 6 columns-per-pore agree with experimental
  pore-to-pore distances within error, but stack too far apart. 6
  column per pore disordered sandwiched and disordered parallel displaced
  configurations stack $\sim$~4.87 and 4.94 \AA~apart respectively, which is
  $\sim$ 1.2~\AA~further apart than suggested by experiment. 5 column-per-pore
  systems stack, at a maximum, 0.9 \AA~further apart than experiment (see
  Table~\ref{table:correlation_length}). The remainder of this discussion will
  therefore focus on the analysis of systems built with 5 columns per pore.

  %BJC: added.
  The number of columns per pore dictates the density of interaction sites within
  each pore and the pore radius. A higher number of interaction sites will 
  surely play a role in transport of molecules that have an affinity for the monomer
  head groups. 
  %We measured the pore radius as the average distance from the pore
  %center to each head group COM. 
  The pore radius increases with the number of 
  columns per pore (See Figure~\ref{S-fig:pore_radius}. This radius is likely 
  related to the size of the molecules which can be excluded by this type of 
  membrane. We may be able to control LLC membrane pore size by selecting monomers
  that are stable in configurations with more columns-per-pore.

  \subsection{Simulated XRD comparison to 2D-WAXS data}

@@

  \subsubsection{The location of R-alkanes}\label{section:ralkanes}

  We normalized the experimental and simulated diffraction patterns so that the
  average intensity within R-alkanes is equal to 1 because we believe that
  R-alkanes is the experimental feature most likely to be reproduced by our
  simulations. We reached this conclusion because alkane parameters in GAFF and
  AMBER forcefields generally reproduce data
  well.~\cite{wang_development_2004,caleman_force_2012}

@@

  \subsubsection{The location and intensity of R-pores}\label{section:rpores}

@@

  \subsubsection{The origin of R-spots}\label{section:rspots}

@@

  Previous literature has attributed the R-spots reflection in this particular
  WAXS dataset as the result of tilted alkane chains~\cite{feng_scalable_2014}.
  This explanation is not unfounded as there are examples in literature of
  systems where tilted liquid crystals give rise to reflections that resemble
  R-spots (Cite, Cite). We measured the tilt angle of the alkane chains of the
  280 K system by measuring the angle made by the vector extending from top to
  bottom of each tail with respect to the membrane plane. We found that it
  equilibrates to an average tilt angle of -2 $\pm$
  13\degree~(Figure~\ref{fig:tilt}), far from the 37\degree~tilt angle previously
  used to explain R-spots. Even when we placed position restraints on the
  monomers in a tilted initial configuration, monomers quickly reduced their tilt
  relative to the $xy$ plane to an angle statisically indistinguishable from zero
  (see Section \ref{S-section:tail_tilt} of the SI). While it is plausible that 
  our force field may in part be responsible for the tendency of monomers to stay
  nearly flat, the monomer tails lack the functional groups which are known to
  facilitate tail tilting.

@@

  \subsubsection{The position, shape, and intensity of R-$\pi$}\label{section:rpi}

@@

  \subsubsection*{Distance and correlation between stacked monomer head groups}

  R-$\pi$ appears at a lower $q_z$ value in our simulations versus experiment
  because monomers in the simulated system stack further apart than those in the
  experimental system (Table~\ref{table:correlation_length}). We calculated
  $z$-direction pair distribution functions, $g(z)$, as described in
  Section~\ref{section:correlation_length}. The resulting distributions are
  generally characterized by decaying oscillatory behavior where the average
  distance between peaks corresponds to the average distance between stacked
  monomer head groups (Figure~\ref{S-fig:correlation}). We calculated this
  equilibrated vertical stacking distance, $\mathit{d}_{equil}$, between monomers
  using Equation~\ref{eqn:decaying_sinusoid}. The distance between stacked
  monomers is greater than experiment by 0.5--0.9 \AA~across all cases, with
  disordered basins at the high end of that range. This behavior is not
  surprising since GAFF models atoms as point charges and does not appropriately
  model the aromatic $\pi-\pi$ interactions, which would make it more
  energetically favorable for the monomers to stack closer together
  \cite{wang_development_2004}. It is also possible that the tails prevent
  close stacking of monomer head groups and we do not achieve the timescales
  necessary for them to rearrange into a necessarily more tightly packed
  configuration. To test this chain-packing hypotheses, we equilibrated the
  system at 500 K and then annealed it to 300 K in an attempt to coerce the tails
  into a more tightly packed configuration. Before annealing and in order to
  stabilize the hexagonal phase at 500 K, we placed harmonic restraints with
  force constants of 1000 kJ mol$^{-1}$ nm$^{-2}$ between head group COMs within
  each column as well as between head group COMs in adjacent columns. We saw no
  improvement in packing distance over our 300 K models (see further discussion
  in Section~\ref{section:slow_dynamics}).

  %BJC: could move the following paragraph to supporting
  The system size, in the $z$-direction, does not significantly alter $g(z)$.
  In Figure~\ref{S-fig:correlation}, oscillations do not fully decay, so a taller
  system may be necessary to fully capture the correlation function. We therefore
  equilibrated a sandwiched system with twice as many monomers per column so that
  the system size doubled in the $z$-direction. The correlation length increases
  modestly from 4.5 $\pm$ 0.4 to 7.3 $\pm$ 1.2 \AA. However, visually, the
  correlation functions are nearly identical and oscillatory behavior persists
  throughout $g(z)$ in both cases (See
  Figure \ref{S-fig:z_correlation_overlay}).

@@

  \subsubsection*{Shape and Intensity of the $q_r$ cross-section of R-$\pi$}

@@

  \subsubsection*{Quenched disorder in ensembles of configurations}

@@

  \subsubsection*{The implications of the new-found understanding of R-$\pi$}

>>>>

  % BJC: reworked text
  \label{rework:purpose}
  A molecular-level understanding of LLC polymer membrane structure, enabled by
  molecular dynamics (MD) simulations, will enhance our comprehension of the role
  of monomer structure in solute transport and provide guidelines to reduce the
  large chemical space available to design monomers for creation of
  separation-specific membranes. It is important that we model the system with a
  nanoscopic pore structure that is maximally consistent with experimental data
  in order to approximate the chemical environment experienced by solutes in 
  future transport studies. Atomistic MD simulations can provide the required
  level of detail (Figure \ref{fig:detailed_pore}), assuming the force fields are
  sufficiently accurate. With such an atomistic model, we can directly observe
  molecular-level solute transport and suggest governing mechanisms. We can also
  observe how the choice of head group interacts with solutes of interest. In
  addition, we can interchange counterions which may influence both the pore size
  and the strength of the Donnan potential.

@@

  We parameterized the interaction potential for the liquid crystal monomer
  Na-GA3C11 using the Generalized AMBER Force Field (GAFF)
  \cite{wang_development_2004} with the Antechamber package
  \cite{wang_automatic_2006} provided with AmberTools16
  \cite{case_ambertools16_2016}. We chose GAFF because it has been parameterized
  for use with organic molecules. Although researchers have optimized force
  fields, including GAFF, for use with liquid crystals, they are often tuned in 
  order to reproduce one or more experimental observables for specific types of
  liquid crystal molecules~\cite{cook_development_2001,boyd_optimization_2015}.
  Since liquid crystals exist in a wide range of shapes and sizes, including candidate
  monomers which we may study, we chose to stick with a widely used force field.
  If we can reproduce experimental trends using a single `out-of-the-box' force
  field, then there will be no need to complicate our studies with customized
  parameter sets. We assigned atomic charges using the am1bccsym method of
  \texttt{molcharge} shipped with QUACPAC from Openeye Scientific Software. We
  ran all molecular dynamics simulations using GROMACS 2016
  \cite{bekker_gromacs:_1993,berendsen_gromacs:_1995,van_der_spoel_gromacs:_2005,hess_gromacs_2008}.

@@

  \subsubsection{\textit{Determining Equilibration Time}}\label{method:equil_time}

@@

  \subsubsection{\textit{Calculation of Pore Spacing}}\label{method:pore_spacing}

@@

  \subsubsection{\textit{Generation of Simulated X-ray Diffraction Patterns}}\label{method:xrd}

@@

  \subsubsection{\textit{Pair Distribution Functions and Correlation Length}}\label{section:correlation_length}

@@

  \subsubsection{\textit{Radial Distribution Functions}}\label{method:rdfs}

@@

  \subsection{\textit{Ionic Conductivity Calculations}}\label{method:ionic_conductivity}

@@

  \subsection{Density of Monomers Around Pores}\label{section:mon_per_pore}

@@

  \label{6colperpore}
  Systems built with 5 columns around each pore equilibrate to a pore spacing
  that is most consistent with the experimental value of 4.12 nm derived from
  SAXS measurements (Figure~\ref{fig:SAXS}). Ordered basin systems built with 4
  columns per pore equilibrate to an average pore spacing 0.25 nm lower than
  experiment. Ordered basin systems built with 6 columns per pore, have an
  equilibrated pore spacing ca. 0.50 nm higher than experiment. Monomers in
  disordered basin systems built with 6 columns-per-pore agree with experimental
  pore-to-pore distances within error, but stack too far apart. 6
  column per pore disordered sandwiched and disordered parallel displaced
  configurations stack $\sim$~4.87 and 4.94 \AA~apart respectively, which is
  $\sim$ 1.2~\AA~further apart than suggested by experiment. 5 column-per-pore
  systems stack, at a maximum, 0.9 \AA~further apart than experiment (see
  Table~\ref{table:correlation_length}). The remainder of this discussion will
  therefore focus on the analysis of systems built with 5 columns per pore.

  %BJC: added.
  \label{addition:pore_radius}
  The number of columns per pore dictates the density of interaction sites within
  each pore and the pore radius. A higher number of interaction sites will 
  surely play a role in transport of molecules that have an affinity for the monomer
  head groups. The pore radius increases with the number of 
  columns per pore (See Figure~\ref{S-fig:pore_radius}). This radius is likely 
  related to the size of the molecules which can be excluded by this type of 
  membrane. One may be able to control LLC membrane pore size by selecting monomers
  that are stable in configurations with more columns-per-pore.

  \subsection{Simulated XRD Comparison to 2D-WAXS Data}

@@

  \subsubsection{The Location of R-alkanes}\label{section:ralkanes}

  We normalized the experimental and simulated diffraction patterns so that the
  average intensity within R-alkanes is equal to 1 because we believe that
  R-alkanes is the experimental feature most likely to be reproduced by our
  simulations. We reached this conclusion because alkane parameters in GAFF and
  AMBER force fields generally reproduce data
  well.~\cite{wang_development_2004,caleman_force_2012}

@@

  \subsubsection{The Location and Intensity of R-pores}\label{section:rpores}

@@

  \subsubsection{The Origin of R-spots}\label{section:rspots}

@@

  \label{addition:rspots}
  Previous literature has attributed the R-spots reflection in this particular
  WAXS dataset as the result of tilted alkane chains~\cite{feng_scalable_2014}.
  This explanation is not unfounded as there are examples in literature of
  systems where tilted liquid crystals give rise to reflections that resemble
  R-spots~\cite{percec_self-assembly_2009,gearba_homeotropic_2007}. We 
  measured the tilt angle of the alkane chains of the 280 K system by measuring
  the angle made by the vector extending from top to
  bottom of each tail with respect to the membrane plane. We found that it
  equilibrates to an average tilt angle of -2 $\pm$
  13\degree~(Figure~\ref{fig:tilt}), far from the 37\degree~tilt angle previously
  used to explain R-spots. Even when we placed position restraints on the
  monomers in a tilted initial configuration, monomers quickly reduced their tilt
  relative to the $xy$ plane to an angle statisically indistinguishable from zero
  (see Section \ref{S-section:tail_tilt} of the SI).

  % There are other mechanism for tilt to occur so the following doesn't say much
  %While it is plausible that 
  %our force field may in part be responsible for the tendency of monomers to stay
  %nearly flat, the monomer tails lack the functional groups which are known to
  %facilitate tail tilting.

@@

  \subsubsection{The Position, Shape, and Intensity of R-$\pi$}\label{section:rpi}

@@

  \subsubsection*{Distance and Correlation Between Stacked Monomer Head Groups}

  R-$\pi$ appears at a lower $q_z$ value in our simulations versus experiment
  because monomers in the simulated system stack further apart than those in the
  experimental system (Table~\ref{table:correlation_length}). We calculated
  $z$-direction pair distribution functions, $g(z)$, as described in
  Section~\ref{section:correlation_length}. The resulting distributions are
  generally characterized by decaying oscillatory behavior where the average
  distance between peaks corresponds to the average distance between stacked
  monomer head groups (Figure~\ref{S-fig:correlation}). We show that the system
  size in the $z$-direction does not significantly alter $g(z)$ in 
  Section~\ref{S-section:correlation_functions} of the SI. We calculated the
  equilibrated vertical stacking distance, $\mathit{d}_{equil}$, between monomers
  using Equation~\ref{eqn:decaying_sinusoid}. The distance between stacked
  monomers is greater than experiment by 0.5--0.9 \AA~across all cases, with
  disordered basins at the high end of that range. This behavior is not
  surprising since GAFF models atoms as point charges and does not appropriately
  model the aromatic $\pi-\pi$ interactions, which would make it more
  energetically favorable for the monomers to stack closer together
  \cite{wang_development_2004}.

% Moved to supporting
%  It is also possible that the tails prevent
%  close stacking of monomer head groups and we do not achieve the timescales
%  necessary for them to rearrange into a necessarily more tightly packed
%  configuration. We tested this hypothesis by annealing systems from high temperatures
%  as outline in Section~\ref{S-chain_packing} of the SI but saw no improvement
%  in packing distance.

%  To test this chain-packing hypotheses, we equilibrated the
%  system at 500 K and then annealed it to 300 K in an attempt to coerce the tails
%  into a more tightly packed configuration. Before annealing and in order to
%  stabilize the hexagonal phase at 500 K, we placed harmonic restraints with
%  force constants of 1000 kJ mol$^{-1}$ nm$^{-2}$ between head group COMs within
%  each column as well as between head group COMs in adjacent columns. We saw no
%  improvement in packing distance over our 300 K models (see further discussion
%  in Section~\ref{section:slow_dynamics}). 
%  
%  %BJC: could move the following paragraph to supporting
%  The system size, in the $z$-direction, does not significantly alter $g(z)$.
%  In Figure~\ref{S-fig:correlation}, oscillations do not fully decay, so a taller
%  system may be necessary to fully capture the correlation function. We therefore
%  equilibrated a sandwiched system with twice as many monomers per column so that
%  the system size doubled in the $z$-direction. The correlation length increases
%  modestly from 4.5 $\pm$ 0.4 to 7.3 $\pm$ 1.2 \AA. However, visually, the
%  correlation functions are nearly identical and oscillatory behavior persists
%  throughout $g(z)$ in both cases (See
%  Figure \ref{S-fig:z_correlation_overlay}).

@@

  \subsubsection*{Shape and Intensity of the $q_r$ Cross-section of R-$\pi$}\label{section:rpi_xsection}

@@

  \subsubsection*{Quenched Disorder in Ensembles of Configurations}

@@

  \subsubsection*{The Implications of the New-found Understanding of R-$\pi$}
